\bf On semidefinite-representable sets over valuation fields

### 2025-05-16 article/paper_padic_spec.tex
commit by Simone

In this section we present an algorithm for solving Problem \eqref{LP}.
The algorithm starts by eliminating the equality constraint defined by $D$ and $e$
by variable substitution. Then the matrix $A$ is replaced by its Smith Normal Form,
and the general linear program is consequently reduced to a diagonal linear program.
The original \eqref{LP} is then easily solved by looking at valuations of input data
in the reduced problem.

We first give a formal description of our algorithm:

\begin{algorithm}
    \caption{\textbf{SOLVELP}($A,b,c,D,e$)}\label{algo_LP}
    \begin{algorithmic}[1]
      \REQUIRE
      $A \in K^{d\times n}$, $b \in K^d$, $c \in K^n$, $D \in K^{m \times n}$ and $e \in K^m$
      \ENSURE [TYPE, $x$], where:
      \vspace{-0.25cm}
      \begin{itemize}
      \item[]
        TYPE $=$ INFEAS and $x=[]$ if and only if \eqref{LP} is infeasible
      \vspace{-0.25cm}
      \item[]
        TYPE $=$ UNBOUND and $x=[]$ if and only if \eqref{LP} is feasible and unbounded
      \vspace{-0.25cm}
      \item[]
        TYPE $=$ FEAS if and only if \eqref{LP} is feasible and has a solution $x$
      \end{itemize}
      \vspace{-0.1cm}
      \hrule
      \vspace{0.1cm}
      \IF{$Dx = e$ has no solution}
      \RETURN [INFEAS, []]
      \ENDIF
      \IF{$D \neq {\bf 0}$}
      \STATE find $J \in K^{n \times R}$, with $R = \dim\ker D$, such that $\text{Im}\, J = \ker D$
      \STATE find $x_0$ such that $Dx_0=e$
      \RETURN $\textbf{SOLVELP}(AJ,b+A x_0,J^T c, {\bf 0},0)$ \hfill {\color{olive} \# \Cref{reducD}}
      \ENDIF
      \IF{$A$ not in Smith Normal Form}
      \STATE $[Q,S,P] \leftarrow \textbf{SNF}(A)$
      \STATE $[\mathrm{STATUS}, y] \leftarrow \textbf{SOLVELP}(S,Qb,(P^{-1})^Tc,{\bf 0},0)$
      \RETURN $[\mathrm{STATUS}, P^{-1}y]$ \hfill {\color{olive} \# \Cref{solsLP}}
      \ELSE
      \IF{$\bigwedge_{i=r+1}^d \val(b_{i}) \geq 0$}
      \STATE STATUS $\leftarrow$ FEAS
      \STATE give solution in this case
      \ENDIF
      \STATE
      \STATE
      \RETURN ... \hfill {\color{olive} \# \Cref{prop:reduc}}
      \ENDIF
    \end{algorithmic}
\end{algorithm}

@@

\begin{lemma}\label{prop:reduc}
  Let $(S,b',c') \in K^{d\times n} \times K^d \times K^n$ be as in \Cref{solsLP}.
  Then:
  \begin{enumerate}
  \item \label{prop:reduc_it1}
    \eqref{LP} with data $(S,b',c',{\bf 0},0)$ is feasible if and only if the $d-r$ last coefficients
    of $b'$ have nonnegative valuation, for $r = {\rm rank}\,S$.
  \item \label{prop:reduc_it2}
    Assume at least one of the $n-r$ last coefficients of $c'$ is non-zero. If \eqref{LP} with
    data $(S,b',c')$ is feasible, then it is unbounded and does not admit solutions.
  \end{enumerate}
\end{lemma}
\begin{proof}
  %%%%\corentin{préciser les notations}. 
  A vector $y = (y_1, \ldots, y_n)^T \in K^n$ is feasible for \eqref{LP} with data
  $(S,b',c',0)$ if and only if $\val(s_i y_i + b'_i) \geq 0$ for\footnote{\simone{J'ai mis
    $\val(s_i y_i + b'_i) \geq 0$ au lieu de $\val(s_i y_i + b_i) \geq 0$ car c'est $b'$
    et non pas $b$.}} all $i = 1,\ldots,n$.
  However $s_i = 0$ for $n-r \leq i\leq n$. Therefore, if \eqref{LP} with data $(S,b',c',0)$ is feasible,
  necessarily the $d-r$ last coefficients of $b'$ have nonnegative valuation.
  Conversely, if the $d-r$ last coefficients of $b'$ are zero, the vector $y \in K^n$ such that
  $y_i = -\frac{-b'_i}{s_i}$ if $1 \leq i \leq r$, and $y_i =0$ otherwise, is feasible for \eqref{LP}
  with data $(S,b',c',{\bf 0},0)$.

>>>>

In this section we present \Cref{algo_LP}, \textbf{SOLVELPval}, for solving Problem \eqref{LP}.
The algorithm starts by eliminating the equality constraint defined by $D$ and $e$
by variable substitution. Then the matrix $A$ is replaced by its Smith Normal Form,
and the general linear program is consequently reduced to a diagonal linear program.
The original \eqref{LP} is then easily solved by looking at valuations of input data
in the reduced problem.

\begin{algorithm}
    \caption{\textbf{SOLVELPval}($A,b,c,D,e$)}\label{algo_LP}
    \begin{algorithmic}[1]
      \REQUIRE
      $A \in K^{d\times n}$, $b \in K^d$, $c \in K^n$, $D \in K^{m \times n}$ and $e \in K^m$
      \ENSURE [TYPE, $x$], where:
      \vspace{-0.25cm}
      \begin{itemize}
      \item[]
        TYPE $=$ INFEAS and $x=[]$ if and only if \eqref{LP} is infeasible
      \vspace{-0.25cm}
      \item[]
        TYPE $=$ UNBOUND and $x=[]$ if and only if \eqref{LP} is feasible and unbounded
      \vspace{-0.25cm}
      \item[]
        TYPE $=$ FEAS if and only if \eqref{LP} is feasible and has a solution $x$
      \end{itemize}
      \vspace{-0.1cm}
      \hrule
      \vspace{0.1cm}
      \IF{$Dx = e$ has no solution}
      \RETURN [INFEAS, []]
      \ENDIF
      \IF{$D \neq {\bf 0}$}
      \STATE find $J \in K^{n \times R}$, with $R = \dim\ker D$, such that $\text{Im}\, J = \ker D$
      \STATE find $x_0$ such that $Dx_0=e$
      \RETURN $\textbf{SOLVELPval}(AJ,b+A x_0,J^T c, {\bf 0},0)$ \hfill {\color{olive} \# \Cref{reducD}}
      \ENDIF
      \IF{$A$ not in Smith Normal Form}
      \STATE $[Q,S,P] \leftarrow \textbf{SNF}(A)$
      \STATE $[\mathrm{STATUS}, y] \leftarrow \textbf{SOLVELPval}(S,Qb,(P^{-1})^Tc,{\bf 0},0)$
      \RETURN $[\mathrm{STATUS}, P^{-1}y]$ \hfill {\color{olive} \# \Cref{solsLP}}
      \ELSE
      \STATE $r \leftarrow \text{rank}(A)$
      \IF{$\bigwedge_{i=r+1}^d \val(b_{i}) \geq 0$}
      \STATE STATUS $\leftarrow$ FEAS
      \STATE give solution in this case
      \ELSE
      \STATE
      \STATE
      \RETURN ... \hfill {\color{olive} \# \Cref{prop:reduc}}
      \ENDIF
      \ENDIF
    \end{algorithmic}
\end{algorithm}

@@

\begin{lemma}\label{prop:reduc}
  Let $(S,b',c') \in K^{d\times n} \times K^d \times K^n$ be as in \Cref{solsLP}.
  Then:
  \begin{enumerate}
  \item \label{prop:reduc_it1}
    \eqref{LP} with data $(S,b',c',{\bf 0},0)$ is feasible if and only if the $d-r$ last coefficients
    of $b'$ have nonnegative valuation, for $r = {\rm rank}\,S$.
  \item \label{prop:reduc_it2}
    Assume at least one of the $n-r$ last coefficients of $c'$ is non-zero. If \eqref{LP} with
    data $(S,b',c')$ is feasible, then it is unbounded and does not admit solutions.
  \end{enumerate}
\end{lemma}
\begin{proof}
  %%%%\corentin{préciser les notations}. 
  A vector $y = (y_1, \ldots, y_n)^T \in K^n$ is feasible for \eqref{LP} with data
  $(S,b',c',0)$ if and only if $\val(s_i y_i + b'_i) \geq 0$ for\footnote{\simone{J'ai mis
    $\val(s_i y_i + b'_i) \geq 0$ au lieu de $\val(s_i y_i + b_i) \geq 0$ car c'est $b'$
    et non pas $b$.}} all $i = 1,\ldots,n$.
  However $s_i = 0$ for $n-r \leq i\leq n$. Therefore, if \eqref{LP} with data $(S,b',c',0)$ is feasible,
  necessarily the $d-r$ last coefficients of $b'$ have nonnegative valuation.
  Conversely, if the $d-r$ last coefficients of $b'$ are zero, the vector $y \in K^n$ such that
  $y_i = {-b'_i}/{s_i}$ if $1 \leq i \leq r$, and $y_i =0$ otherwise, is feasible for \eqref{LP}
  with data $(S,b',c',{\bf 0},0)$.
